Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-05-30 tacl.tex
Pushed it under 13 pages

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence:
\begin{displaymath}
\minLPR(w_{1,n}) = \min_{1 \leq i \leq n }{\LPR(w_i,w_{1,n})}
\end{displaymath}
Here, minLPR for a particular $n$-gram does not reflect the \emph{overall} degree to which it holds together, but rather focuses on the word which is its weakest link. For example, in the case of \ex{be keep \gap under wraps}, a general statistical metric might assign it a high score due to the strong association between \ex{keep} and \ex{under} or \ex{under} and \ex{wraps}, but minLPR is focused on the weaker relationship between \ex{be} and the rest of the phrase. This makes it particularly suited to use in a lattice model of competing $n$-grams, where the choice of \ex{be keep \gap under wraps} versus \ex{keep \gap under wraps} should be based exactly on the extent to which \ex{be} is an essential part of the phrase; the other affinities are, in effect, irrelevant, because they occur in the smaller $n$-gram as well.

@@

In order to capture the current state of the model and provide an overall explainedness score, we introduce two indicator functions: $I_{\mathit{on}}(t)$ is 1 if $t$ is on and 0 if $t$ is off ; $I_{\mathit{off}}(t)$ is the reverse. The objective function maximized by the model is then the explainedness ($\mathit{expl}$) across all the nodes of the lattice, $t_1,\dots,t_{|T|} \in T$, which can be defined in terms of minLPR and the node interaction functions:
%
\begin{multline}
	\mathit{expl}(T) = \prod_{i=1}^{|T|}{I_{\mathit{on}}(t_i)\cdot C^{-\overlap(t_1)}} \\ + I_{\mathit{off}}(t_i)\cdot\minLPR(t_i)^{-\cover(t_i) \cdot \clear(t_i)}
\end{multline}
\noindent
When a node is off, its explainedness is the inverse of its minLPR, except if there are covering or clearing nodes which explain it by pushing the exponent of minLPR towards zero. When the node is on, its explainedness is the inverse of a fixed cost hyperparameter $C$, though this cost is increased if it overlaps with other active nodes. 
All else being equal, when $\minLPR(t) > C$, a node will be selected as an FS, and so, independent of the node interactions, $C$ can be viewed as the threshold for the minLPR association measure under a traditional approach to MWE identification. There is no upper bound on $C$, but empirically, we have found values in the range $[3,6]$ give reasonable results for the languages presented in this paper.

@@

Other than the inclusion of new languages, our test sets differ from \newcite{Brooke15b} in two ways. One advantage of a type-based annotation approach, particularly with regards to annotation with a known subjective component, is that it is quite sensible to simply discard borderline cases, improving reliability at the cost of some representativeness. To this end, we entirely excluded from our test set $n$-grams which just one annotator marked as FS. \tabref{tab:stats} contains the counts for the four test sets after this filtering step as well as Cohen's Kappa scores before (Pre) and after (Post).  The second change is that for the main evaluation we collapsed gapped and contiguous $n$-grams into a single test set. The rationale is that the number of positive gapped examples is too low to provide a reliable independent F-score (see further discussion in \secref{sec:discussion}).

@@

	Lattice \emph{$-$ovr} & 0.52& 0.53 & 0.53 & & 0.34& \bf{0.60} & 0.44& & 0.36 & \bf{0.67} & 0.47 & & 0.53 & \bf{0.62} & 0.58 \\

@@

 \end{tabular}
 \caption{Results of FS identification in various test sets: Countrank = ranking with frequency; PMIrank = PMI-based ranking; minLPRrank = ranking with minLPR; LPRseg = the method of Brooke et \al (2015); ``$-$cl'' = no clearing; ``$-$ovr'' = no penalization of overlaps; ``P'' = Precision; ``R'' = Recall; and ``F'' = F-score. Bold is best in a given column. For ICWSM only, F-score difference relative to best baseline is significant at $p < 0.05$; the difference for all test sets together is significant at $ p < 0.01$ (all $p$ values based on the permutation test).}
	 \label{tab:main}

@@

We created small development sets for each corpus and used them to do a thorough testing of parameter settings. Although it is generally possible to increase precision by increasing $C$, we found that across corpora we always obtained near-optimal results with $C=4$, so to demonstrate the usefulness of the lattice technique as an entirely off-the-shelf tool, we present the results using identical settings for all four corpora. We treat covering as a fundamental part of the Lattice model, but to investigate the efficacy of other node interactions within the model we present results with each of overlap and clearing node interactions turned off.

@@

The main results for FS acquisition across the four corpora are shown in \tabref{tab:main}. As noted in \secref{sec:background}, simple statistical association measures like PMI do poorly when faced with syntactically-unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of \newcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across the languages. Generally, clearing and overlap node interactions provide a relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian.  When only covering is used, the results are very similar to \newcite{Brooke15b}, which is unsurprising given the extent to which decomposition and covering are related. The Japanese and ICWSM corpus have relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

@@

Though the results across the four corpora are reasonably similar with respect to overall F-score, there are some discrepancies. By using the standard UNIDIC morpheme representation as the base unit for Japanese, the model ends up doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are relatively more FS for Japanese: precision is high, and recall is comparably low.  Importantly, the initial $n$-gram statistics derived from corpus actually reflect that Japanese is different: the number of $n$-gram types over length 4 is almost twice the number in the ICWSM corpus. One idea for future work is to automatically adapt to the input language/corpus in order to ensure a good balance between precision and recall.

@@

The statistics in \tabref{tab:lexstats} otherwise reflect the quantity and diversity of FS across the corpora, particularly in terms of the number of POS patterns represented in the lexicon. Looking at the most common POS patterns across languages, only noun-noun and adjective-noun combinations ever account for more than 5\% of all word types in any of the lexicons. Though some of the diversity can of course be attributed to noise, it is safe to say that most FS do not fall into the standard two-word syntactic categories used in MWE work, and therefore identifying them requires a much more general approach, like the one presented here.

@@

%The fact that less than 20\% of the FS derived from the BNC appear in the lexicon for the ICWSM corpus is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.
\tabref{tab:sample} contains 10 randomly selected examples from each of the corpora. Among the English examples, most of the clear errors are bigrams that reflect particular biases of their respective corpora: The phrase \ex{via slashdot} comes from boilerplate text identifying the source of a article, whereas \ex{Maureen} (from \ex{Maureen says}) is a character in one of the novels included in the BNC. The longer FS mostly seem sensible, in that they are plausible lexicalized constructions, though \ex{primarily as \gap and} from the BNC is not obviously an FS and likely the result of noise. Some FS are dialectal variants, for instance \ex{license endorsed} refers to the British traffic violations. More generally, the FS lexicons created by these two corpora are quite distinct.

What is perhaps most striking about the non-English examples is how poorly some FS translate: many good FS in these languages become extremely awkward when translated into English. This is expected of course for idioms like \ex{biti general poslije bitke} \gl{be the general after the battle} (\ie \gl{hindsight is 20/20}), but it extends to other relatively compositional  constructions like 
\begin{CJK*}{UTF8}{min} \ex{こう\ 言う\ \gap\ が\ 続く} \end{CJK*} 
\gl{repeat occurrences of \gap like this} and 
\begin{CJK*}{UTF8}{min} \ex{前期\ 比} \end{CJK*} 
\gl{first half comparison}. This highlights the importance of focusing on FS when learning a language, since many FS do not translate well (though some short ones do, particularly for languages in the same family). Though some of the errors seem to be the result of extra material added to an good FS, for instance \ex{promet teretnih vozila} \gl{good vehicle traffic}, most, again, are somewhat inexplicable artifacts of the corpus they were built from collected, like \ex{austrijski investitor} \gl{Austrian investor}.

Since Zipfian frequency curves naturally extend to multiword FS vocabulary, our lexicons (and type-based evaluation of the them) are of course dominated by rarer terms. This is not, we would argue, a serious drawback, since in practical terms there is very little value in focusing on common MWEs like \ex{of course} which manually-built lexicons will cover; most of potential in automatic extraction comes from the long tail. However, we did investigate the other end of the Zipfian curve by extracting the 20 most common MWEs (including both strong and weak) from the \newcite{Schneider14a} corpus. In the ICWSM lexicon, our recall for these common terms was fairly high (0.75), with errors mostly resulting from longer phrases containing these terms ``winning out'' (in the lattice) over shorter phrases, which have relatively low LPR due to extremely common constituent words; for example, we missed \ex{on time}, but had 19 FS which contain it (e.g.\ \ex{right on time}, \ex{show up on time}, and \ex{start on time}). In one case which showed this same problem, \ex{waste * time}, the lexicon did have its ungapped version, highlighting potential for improvement.

@@

In \secref{sec:background}, we noted than FS is generally a much broader category than MWE, which we take as referring to terms which carry significant non-compositional meaning. We decided to investigate the distinction at a practical level by annotating the positive examples in the ICWSM test set for being MWE or non-MWE FS.\footnote{The set was exhaustively annotated by two native-speaker annotators ($\kappa = 0.73$), and conflicts were resolved through discussion.} First, we note that only 28\% of our FS types were labeled MWE; this is in contrast to, for instance, the annotation of \newcite{Schneider14a} where ``weak'' MWE make up a small fraction of MWE types. Even without any explicit representation of compositionality, our model did much better at identifying MWE FS than non-MWE FS: recall for the former was 0.71, whereas recall for the latter was just 0.34. This may simply reflect, however, the fact that a disproportionate number of MWEs were noun-noun compounds, which are fairly easy for the model to identify.

Due to the lack of spaces between words and an agglutinative morphology, the standard approach to tokenization and lemmatization in Japanese involves morphological rather than word segmentation. In terms of the content of the resulting lexicon we believe the effect of this difference is modest, since much of the extra FS in Japanese would simply be individual words in other languages (which could be viewed as being trivially part of the FS lexicon). However, from a theoretical perspective we might very much prefer to build FS for all languages starting from morphemes rather than words. Such a framework could, for instance, capture inflectional flexibilty versus fixedness directly in the model, with fixed inflectional morphemes included as a distinct element of the FS and flexible morphemes becoming gaps. However, for many languages this would result in a huge blow up in complexity with only modest increases in the scope of FS identification. Though it is indisputable that inflectional fixedness is part of the lexical information contained in an FS, in practice this sort of information can be efficiently derived post hoc using corpus statistics.

Though we have demonstrated that competition within a lattice is powerful for the production of FS lexicons, its usefulness derives less from the specific choices we have made in this instantiation of the model, and more from the flexiblity that such a model provides for future research. Not only do alternatives like DP-seg and LocalMaxs fail to scale up to large corpora, there are few obvious ways to improve on their simple underlying algorithms without compromising their elegance and worsening tractability. Fast and functional, the LPR decomp approach is nevertheless algorithmically ungainly, involving multiple layers of heuristic-driven filtering with no possibility of correcting errors. Our lattice method is aimed at something between these extremes: a practical, optimizable model, but with various component heuristics that can be improved upon. For instance, though the current version of clearing is effective and has practical advantages relative to the other simpler options that we tested, we think that it could be enhanced by more careful investigation of the statistical properties of FS, identifying to what extent the LPR of words around FS is likely to be inflated.

We can also consider adding new terms to the exponents of the two terms of our objective function, analagous to the cover, clear, and overlap functions, based on other relationships between nodes in the lattice. One which we have considered is creating new connections between identical or similar syntactic patterns, which could serve to encourage the model to generalize. In English, for instance, it might learn that verb-particle combinations are generally likely to be FS, whereas verb-determiner combinations are not. Our initial investigations suggest, however, it may be difficult to apply this idea without merely amplifying existing undesirable biases in the LPR measure. Bringing in other information such as simple distributional statistics might help the model identify non-compositional semantics, and could, in combination with the existing lattice competition, focus the model on detecting MWEs, which could in turn serve as a reliable basis for generalization when the raw LPR statistics are an unreliable indicator of formulaticity.

With respect to speed, although the optimization of the lattice is several orders of magnitude slower than the decomposition heuristics of \newcite{Brooke15b}, the time needed to build and optimize the lattice is still only a fraction of the time required to collect the statistics necessary to calculate LPR: therefore, the overall effect on runtime of the lattice optimization step is fairly minimal. In the BNC, the end-to-end lattice algorithm is several times faster than LocalMaxs (with a fraction of its memory usage), and an order of magnitude faster than DP-Seg. We note that for all four corpora, the lattice optimization algorithm converged within 10 iterations.

@@

Finally, though the model presented here was designed specifically for FS extraction, we note that it could be useful for related tasks such as unsupervised learning of morphological lexicons, particularly for agglutinative languages. Character or phoneme $n$-grams could compete in an identically structured lattice (though presumably without gapped $n$-grams) to be chosen as the best morphemes for the language, and we could straightforwardly adapt LPR to use phonological predictability (\ie based on vowel/consonant ``tags'') instead of syntactic predictability. It is likely, though, that further algorithmic modifications would be necessary to target morphological phenomena well, and we leave this for future work.

@@

>>>>

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence:
\begin{displaymath}
\minLPR(w_{1,n}) = \min_{1 \leq i \leq n }{\LPR(w_i,w_{1,n})}
\end{displaymath}
Here, minLPR for a particular $n$-gram does not reflect the \emph{overall} degree to which it holds together, but rather focuses on the word which is its weakest link. For example, in the case of \ex{be keep \gap under wraps} (\figref{fig:example}), a general statistical metric might assign it a high score due to the strong association between \ex{keep} and \ex{under} or \ex{under} and \ex{wraps}, but minLPR is focused on the weaker relationship between \ex{be} and the rest of the phrase. This makes it particularly suited to use in a lattice model of competing $n$-grams, where the choice of \ex{be keep \gap under wraps} versus \ex{keep \gap under wraps} should be based exactly on the extent to which \ex{be} is an essential part of the phrase; the other affinities are, in effect, irrelevant, because they occur in the smaller $n$-gram as well.

@@

In order to capture the current state of the model and provide an overall explainedness score, we introduce two indicator functions: $I_{\mathit{on}}(t)$ is 1 if node $t$ is on and 0 if $t$ is off ; $I_{\mathit{off}}(t)$ is the reverse. The objective function maximized by the model is then the explainedness ($\mathit{expl}$) across all the nodes of the lattice, $t_1,\dots,t_{|T|} \in T$, which can be defined in terms of minLPR and the node interaction functions:
%
\begin{multline}
	\mathit{expl}(T) = \prod_{i=1}^{|T|}{I_{\mathit{on}}(t_i)\cdot C^{-\overlap(t_1)}} \\ + I_{\mathit{off}}(t_i)\cdot\minLPR(t_i)^{-\cover(t_i) \cdot \clear(t_i)}
\end{multline}
\noindent
When a node is off, its explainedness is the inverse of its minLPR, except if there are covering or clearing nodes which explain it by pushing the exponent of minLPR towards zero. When the node is on, its explainedness is the inverse of a fixed cost hyperparameter $C$, though this cost is increased if it overlaps with other active nodes. 
All else being equal, when $\minLPR(t) > C$, a node will be selected as an FS, and so, independent of the node interactions, $C$ can be viewed as the threshold for the minLPR association measure under a traditional approach to MWE identification. There is no upper bound on $C$, but empirically, we have found values in the range $[3,6]$ give reasonable results for the languages presented in this paper.

@@

Other than the inclusion of new languages, our test sets differ from \newcite{Brooke15b} in two ways. One advantage of a type-based annotation approach, particularly with regards to annotation with a known subjective component, is that it is quite sensible to simply discard borderline cases, improving reliability at the cost of some representativeness. To this end, we entirely excluded from our test set $n$-grams which just one annotator marked as FS. \tabref{tab:stats} contains the counts for the four test sets after this filtering step as well as Cohen's Kappa scores before (Pre) and after (Post).  The second change is that for the main evaluation we collapsed gapped and contiguous $n$-grams into a single test set. The rationale is that the number of positive gapped examples is too low to provide a reliable independent F-score.

%(see further discussion in \secref{sec:discussion}).

@@

	Lattice \emph{$-$ovr} & 0.52& \bf{0.53} & 0.53 & & 0.34& \bf{0.60} & 0.44& & 0.36 & \bf{0.67} & 0.47 & & 0.53 & \bf{0.62} & 0.58 \\

@@

 \end{tabular}
 \caption{Results of FS identification in various test sets: Countrank = ranking with frequency; PMIrank = PMI-based ranking; minLPRrank = ranking with minLPR; LPRseg = the method of Brooke et \al (2015); ``$-$cl'' = no clearing; ``$-$ovr'' = no penalization of overlaps; ``P'' = Precision; ``R'' = Recall; and ``F'' = F-score. Bold is best in a given column. For ICWSM only, F-score difference relative to best baseline is significant at $p < 0.05$; the difference for all test sets together is significant at $ p < 0.01$; all $p$ values based on the permutation test (Yeh, 2000).}
	 \label{tab:main}

@@

We created small development sets for each corpus and used them to do a thorough testing of parameter settings. Although it is generally possible to increase precision by increasing $C$, we found that across corpora we always obtained near-optimal results with $C=4$, so to demonstrate the usefulness of the lattice technique as an entirely off-the-shelf tool, we present the results using identical settings for all four corpora. We treat covering as a fundamental part of the Lattice model, but to investigate the efficacy of other node interactions within the model we present results with overlap and clearing node interactions turned off.

@@

The main results for FS acquisition across the four corpora are shown in \tabref{tab:main}. As noted in \secref{sec:background}, simple statistical association measures like PMI do poorly when faced with syntactically-unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of \newcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across the languages. Generally, clearing and overlap node interactions provide a relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian.  When only covering is used, the results are fairly similar to \newcite{Brooke15b}, which is unsurprising given the extent to which decomposition and covering are related. The Japanese and ICWSM corpus have relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

@@

Though the results across the four corpora are reasonably similar with respect to overall F-score, there are some discrepancies. By using the standard UNIDIC morpheme representation as the base unit for Japanese, the model ends up doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are relatively more FS for Japanese: precision is high, and recall is comparably low.  Importantly, the initial $n$-gram statistics actually reflect that Japanese is different: the number of $n$-gram types over length 4 is almost twice the number in the ICWSM corpus. One idea for future work is to automatically adapt to the input language/corpus in order to ensure a good balance between precision and recall.

@@

The statistics in \tabref{tab:lexstats} otherwise reflect the quantity and diversity of FS across the corpora, particularly in terms of the number of POS patterns represented in the lexicon. Looking at the most common POS patterns across languages, only noun-noun and adjective-noun combinations ever account for more than 5\% of all word types in any of the lexicons. Though some of the diversity can of course be attributed to noise, it is safe to say that most FS do not fall into the standard two-word syntactic categories used in MWE work, and therefore identifying them requires a much more general approach like the one presented here.

@@

%The fact that less than 20\% of the FS derived from the BNC appear in the lexicon for the ICWSM corpus is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.
\tabref{tab:sample} contains 10 randomly selected examples from each of the lexicons produced by our method. Among the English examples, most of the clear errors are bigrams that reflect particular biases of their respective corpora: The phrase \ex{via slashdot} comes from boilerplate text identifying the source of a article, whereas \ex{Maureen} (from \ex{Maureen says}) is a character in one of the novels included in the BNC. The longer FS mostly seem sensible, in that they are plausible lexicalized constructions, though \ex{primarily as \gap and} from the BNC is not obviously an FS and likely the result of noise. Some FS are dialectal variants, for instance \ex{license endorsed} refers to the British traffic violations. More generally, the FS lexicons created by these two corpora are quite distinct, sharing less than 50\% of their entries.

What is perhaps most striking about the non-English examples is how poorly some FS translate: many good FS in these languages become extremely awkward when translated into English. This is expected of course for idioms like \ex{biti general poslije bitke} \gl{be the general after the battle} (\ie \gl{hindsight is 20/20}), but it extends to other relatively compositional  constructions like 
\begin{CJK*}{UTF8}{min} \ex{こう\ 言う\ \gap\ が\ 続く} \end{CJK*} \gl{repeat occurrences of \gap like this} and 
\begin{CJK*}{UTF8}{min} \ex{前期\ 比} \end{CJK*} 
\gl{first half comparison}. This highlights the importance of focusing on FS when learning a language, since many FS do not translate well (though some do, particularly short ones in related languages). Though some of the errors seem to be the result of extra material added to an good FS, for instance \ex{promet teretnih vozila} \gl{good vehicle traffic}, most, again, are somewhat inexplicable artifacts of the corpus they were built from, like \ex{austrijski investitor} \gl{Austrian investor}.

Since Zipfian frequency curves naturally extend to multiword vocabulary, our lexicons (and type-based evaluation of the them) are of course dominated by rarer terms. This is not, we would argue, a serious drawback, since in practical terms there is very little value in focusing on common FS like \ex{of course} which manually-built lexicons already contain; most of potential in automatic extraction comes from the long tail. However, we did investigate the other end of the Zipfian curve by extracting the 20 most common MWEs (including both strong and weak) from the \newcite{Schneider14a} corpus. In the ICWSM lexicon, our recall for these common terms was fairly high (0.75), with errors mostly resulting from longer phrases containing these terms ``winning out'' (in the lattice) over shorter phrases, which have relatively low LPR due to extremely common constituent words; for example, we missed \ex{on time}, but had 19 FS which contain it (e.g.\ \ex{right on time}, \ex{show up on time}, and \ex{start on time}). 
%In one case which showed this same problem, \ex{waste * time}, the lexicon did have its ungapped version, highlighting potential for improvement.

@@

In \secref{sec:background}, we noted than FS is generally a much broader category than MWE, which we take as referring to terms which carry significant non-compositional meaning. We decided to investigate the distinction at a practical level by annotating the positive examples in the ICWSM test set for being MWE or non-MWE FS.\footnote{The set was exhaustively annotated by two native-speaker annotators ($\kappa = 0.73$), and conflicts were resolved through discussion.} First, we note that only 28\% of our FS types were labeled MWE; this is in contrast to, for instance, the annotation of \newcite{Schneider14a} where ``weak'' MWE make up a small fraction of MWE types. Even without any explicit representation of compositionality, our model did much better at identifying MWE FS than non-MWE FS:  0.71 versus 0.34 recall. This may simply reflect, however, the fact that a disproportionate number of MWEs were noun-noun compounds, which are fairly easy for the model to identify.

Due to the lack of spaces between words and an agglutinative morphology, the standard approach to tokenization and lemmatization in Japanese involves morphological rather than word segmentation. In terms of the content of the resulting lexicon we believe the effect of this difference on FS extraction is modest, since much of the extra FS in Japanese would simply be single words in other languages (and considered trivially part of the FS lexicon). However, from a theoretical perspective we might very much prefer to build FS for all languages starting from morphemes rather than words. Such a framework could, for instance, capture inflectional flexibilty versus fixedness directly in the model, with fixed inflectional morphemes included as a distinct element of the FS and flexible morphemes becoming gaps. However, for many languages this would result in a huge blow up in complexity with only modest increases in the scope of FS identification. Though it is indisputable that inflectional fixedness is part of the lexical information contained in an FS, in practice this sort of information can be efficiently derived post hoc from the corpus statistics.

Though we have demonstrated that competition within a lattice is a powerful method for the production of FS lexicons, its usefulness derives less from the specific choices we have made in this instantiation of the model, and more from the flexiblity that such a model provides for future research. Not only do alternatives like DP-seg and LocalMaxs fail to scale up to large corpora, there are few obvious ways to improve on their simple underlying algorithms without compromising their elegance and worsening tractability. Fast and functional, the LPR decomp approach is nevertheless algorithmically ungainly, involving multiple layers of heuristic-driven filtering with no possibility of correcting errors. Our lattice method is aimed at something between these extremes: a practical, optimizable model, but with various component heuristics that can be improved upon. For instance, though the current version of clearing is effective and has practical advantages relative to simpler options that we tested, it could be enhanced by more careful investigation of the statistical properties of $n$-grams which contain FS.

We can also consider adding new terms to the exponents of the two parts of our objective function, analagous to the cover, clear, and overlap functions, based on other relationships between nodes in the lattice. One which we have considered is creating new connections between identical or similar syntactic patterns, which could serve to encourage the model to generalize. In English, for instance, it might learn that verb-particle combinations are generally likely to be FS, whereas verb-determiner combinations are not. Our initial investigations suggest, however, it may be difficult to apply this idea without merely amplifying existing undesirable biases in the LPR measure. Bringing in other information such as simple distributional statistics might help the model identify non-compositional semantics, and could, in combination with the existing lattice competition, focus the model on MWEs which could provide a reliable basis for generalization.

For all four corpora, the lattice optimization algorithm converged within 10 iterations.  Although the optimization of the lattice is nevertheless several orders of magnitude slower than the decomposition heuristics of \newcite{Brooke15b}, the time needed to build and optimize the lattice is still only a fraction of the time required to collect the statistics for LPR calculation, and so the end-to-end runtime of the two methods is comparable. In the BNC, the end-to-end lattice method is several times faster than LocalMaxs (with a fraction of its memory usage), and an order of magnitude faster than DP-Seg.

@@

Finally, though the model presented here was designed specifically for FS extraction, we note that it could be useful for related tasks such as unsupervised learning of morphological lexicons, particularly for agglutinative languages. Character or phoneme $n$-grams could compete in an identically structured lattice to be chosen as the best morphemes for the language, with LPR adapted to use phonological predictability (\ie based on vowel/consonant ``tags'') instead of syntactic predictability. It is likely, though, that further algorithmic modifications would be necessary to target morphological phenomena well, and we leave this for future work.

@@

\nocite{yeh2000more}
